---
title: "COVID-19"
author: "Aimee Ouyang"
date: "8/7/2021"
output: html_document
---

### import data 

```{r}
library(tidyverse)
covid <- read.csv("covidDataOutput.csv")    # load files
covid <- covid[,c(2:3,6,8)]    # select deaths and deaths
covid$date <- gsub("2021-([0-9,-]+) .*", "\\1", covid$date)
covid <- as.data.frame(t(covid))    # change col and row
names(covid) <- as.matrix(covid[2, ])    #set the time line to header
covid <- as.data.frame(t(covid))    # change col and row
covid[] <- lapply(covid, function(x) type.convert(as.character(x)))
```

```{r}
# create a table display deaths information
deaths <- covid[,c(1:2,4)]
first <- deaths[1:31,]
for (i in (1:(nrow(deaths)/31))) {
 second <- deaths[(i*31+1):((i+1)*31),]
 first <- cbind(first, second)
} 
selected <- seq(3,605,3)
deaths <- data.frame(x=first[,selected])
names(deaths) <- c("AFG", "AGO", "ALB", "AND", "ARE", "ARG", "ARM", "ASM", "ATG", "AUS", "AUT", "AZE", "BDI", "BEL", "BEN", "BFA", "BGD", "BGR", "BHR", "BHS", "BIH", "BLR", "BLZ", "BMU", "BOL", "BRA", "BRB", "BRN", "BTN", "BWA", "CAC", "CAF", "CAN", "CHE", "CHL", "CHN", "CIV", "CMR", "COD", "COG", "COL", "COM", "CPV", "CRI", "CUB", "CYP", "CZE", "DEU", "DJI", "DMA", "DNK", "DOM", "DPC", "DZA", "ECU", "EGY", "ERI", "ESP", "EST", "ETH", "FIN", "FJI", "FRA", "GAB", "GBR", "GEO", "GHA", "GIN", "GMB", "GNB", "GNQ", "GPC", "GRC", "GRD", "GTM", "GUM", "GUY", "HND", "HRV", "HTI", "HUN", "IDN", "IND", "IRL", "IRN", "IRQ", "ISL", "ISR", "ITA", "JAM", "JOR", "JPN", "KAZ", "KEN", "KGZ", "KHM", "KIR", "KNA", "KOR", "KWT", "LAO", "LBN", "LBR", "LBY", "LCA", "LIE", "LKA", "LSO", "LTU", "LUX", "LVA", "MAR", "MCO", "MDA", "MDG", "MDV", "MEX", "MHL", "MKD", "MLI", "MLT", "MMR", "MNE", "MNG", "MNP", "MOZ", "MRT", "MSZ", "MUS", "MWI", "MYS", "NAM", "NER", "NGA", "NIC", "NLD", "NOR", "NPL", "NZL", "OMN", "PAK", "PAN", "PER", "PHL", "PLW", "PNG", "POL", "PRI", "PRT", "PRY", "PSE", "QAT", "RKS", "ROU", "RUS", "RWA", "SAU", "SDN", "SEN", "SGP", "SLB", "SLE", "SLV", "SMR", "SOM", "SRB", "SSD", "STP", "SUR", "SVK", "SVN", "SWE", "SWZ", "SYC", "SYR", "TCD", "TGO", "THA", "TJK", "TLS", "TTO", "TUN", "TUR", "TWN", "TZA", "UGA", "UKR", "URY", "USA", "UZB", "VAT", "VCT", "VEN", "VIR", "VNM", "VUT", "WSM", "YEM", "ZAF", "ZMB","ZWE")
deaths <- as.data.frame(t(deaths))    # change col and row
deaths <- deaths[complete.cases(deaths),]    # remove all the NAs
names(deaths) <- c("X01", "X02", "X03", "X04", "X05", "X06", "X07", "X08", "X09", "X10", "X11", "X12", "X13", "X14", "X15", "X16", "X17", "X18", "X19", "X20", "X21", "X22", "X23", "X24", "X25", "X26", "X27", "X28", "X29", "X30", "X31")
```

```{r}
# Random shuffle data, and set 70% train 30% test. 
set.seed(58)
rows <- sample(nrow(deaths))
deaths <- deaths[rows, ]

deaths_train <- deaths[1:131,]
deaths_test <- deaths[132:187,]
```

### Random Forests

```{r include=FALSE}
library(randomForest)
```

```{r}
# create a random forest with 100 trees, at most 10 nodes in the tree, and consider 5 vaiables in each node. 
set.seed(58)
rf_model <- randomForest(X31 ~ .
                    , data = deaths_train
                    , mtry = 5
                    , maxnods = 10
                    , ntree = 100 
                    , importance = TRUE
                    , type = "prob")
rf_model
```

```{r}
# find the importance of attributes
varImpPlot(rf_model)
```

```{r}
# prediction result
deaths_test$pred_rf <- predict(rf_model, newdata = deaths_test)
```

```{r}
library(Metrics)    # library for using rmse()
rf_rmse <- rmse(deaths_test$X31, deaths_test$pred_rf)
```


### Gradient Boosting

```{r}
# Helper packages
library(dplyr)    # for general data wrangling needs

# Modeling packages
library(gbm)      # for original implementation of regular and stochastic GBMs
library(h2o)      # for a java-based implementation of GBM variants
library(xgboost)  # for fitting extreme gradient boosting
```

```{r}
# create a gradient boosting with specified shrinkage and n.trees evluated below
# After evluation, we choose 4950 number of trees and 0.1 learning rate (shrinkage) with 10 cross-validation, at least 7 obervations, and 5 number of splits in each terminal node. 
set.seed(58)  # for reproducibility
ames_gbm1 <- gbm(
  formula = X31 ~ .,
  data = deaths_train,
  distribution = "gaussian",  # SSE loss function
  n.trees = 4950,
  shrinkage = 0.1,
  interaction.depth = 5,
  n.minobsinnode = 5,
  cv.folds = 10
)

# find index for number trees with minimum CV error
best <- which.min(ames_gbm1$cv.error)

# get MSE and compute RMSE
gb_rmse <- sqrt(ames_gbm1$cv.error[best])
```

```{r}
# create grid search
hyper_grid <- expand.grid(
  learning_rate = c(0.3, 0.1, 0.05, 0.01, 0.005),
  RMSE = NA,
  trees = NA,
  time = NA
)

# execute grid search
for(i in seq_len(nrow(hyper_grid))) {

  # fit gbm
  set.seed(123)  # for reproducibility
  train_time <- system.time({
    m <- gbm(
      formula = X31 ~ .,
      data = deaths_train,
      distribution = "gaussian",
      n.trees = 5000, 
      shrinkage = hyper_grid$learning_rate[i], 
      interaction.depth = 3, 
      n.minobsinnode = 10,
      cv.folds = 10 
   )
  })
  
  # add SSE, trees, and training time to results
  hyper_grid$RMSE[i]  <- sqrt(min(m$cv.error))
  hyper_grid$trees[i] <- which.min(m$cv.error)
  hyper_grid$Time[i]  <- train_time[["elapsed"]]

}

# results
arrange(hyper_grid, RMSE)
```
A learning rate of 0.100 sufficiently minimizes our loss function and requires 642 trees

```{r}
# search grid
hyper_grid <- expand.grid(
  n.trees = 642,
  shrinkage = 0.100,
  interaction.depth = c(3, 5, 7),
  n.minobsinnode = c(5, 10, 15)
)

# create model fit function
model_fit <- function(n.trees, shrinkage, interaction.depth, n.minobsinnode) {
  set.seed(123)
  m <- gbm(
    formula = X31 ~ .,
    data = deaths_train,
    distribution = "gaussian",
    n.trees = n.trees,
    shrinkage = shrinkage,
    interaction.depth = interaction.depth,
    n.minobsinnode = n.minobsinnode,
    cv.folds = 10
  )
  # compute RMSE
  sqrt(min(m$cv.error))
}

# perform search grid with functional programming
hyper_grid$rmse <- purrr::pmap_dbl(
  hyper_grid,
  ~ model_fit(
    n.trees = ..1,
    shrinkage = ..2,
    interaction.depth = ..3,
    n.minobsinnode = ..4
    )
)

# results
arrange(hyper_grid, rmse)
```
When the interaction.depth = 5 and n.minobsinode = 5, we obtained the smallest rmse. 

```{r}
# find the importance of attribute used in this model
summary(m)    
```

```{r}
n.trees = seq(from=650 ,to=5000, by=50) #no of trees-a vector of 100 values 

#Generating a Prediction matrix for each Tree
predmatrix<-predict(m,deaths_test,n.trees = n.trees)
dim(predmatrix) #dimentions of the Prediction Matrix

#Calculating The Mean squared Test Error
test.error<-with(deaths_test,apply( (predmatrix-X31)^2,2,mean))
head(test.error) #contains the Mean squared test error for each of the 100 trees averaged

#Plotting the test error vs number of trees

plot(n.trees , test.error , pch=19,col="blue",xlab="Number of Trees",ylab="Test Error", main = "Perfomance of Boosting on Test Set")

#adding the RandomForests Minimum Error line trained on same data and similar parameters
abline(h = min(test.error),col="red") #test.err is the test error of a Random forest fitted on same data
legend("topright",c("Minimum Test error Line for Gradient Boosting"),col="red",lty=1,lwd=1)
```
From the graph, using 400 decision trees provides the lowest mean squared test error. 

### K-Nearest Neighbor Regression (KNN) 

```{r}
library('class')
```

```{r}
# build a k-NN model with 6 nearest neighbors and make predictions
set.seed(58)
deaths_test$kn <- knn(deaths_train[,1:30], deaths_test[,1:30], deaths_train$X31, k=6)
deaths_test$kn <- as.character(deaths_test$kn)
deaths_test$kn <- as.numeric(deaths_test$kn)
kn_rmse <- rmse(deaths_test$X31, deaths_test$kn)
```

### LASSO

```{r}
# load packages
library(data.table) # used for reading and manipulation of data
library(dplyr)	 # used for data manipulation and joining
library(glmnet)	 # used for regression
library(ggplot2) # used for ploting
library(caret)	 # used for modeling
library(xgboost) # used for building XGBoost model
library(e1071)	 # used for skewness
library(cowplot) # used for combining multiple plots
```

```{r}
# Model Building :Lasso Regression
set.seed(123)
control = trainControl(method ="cv", number = 5)
Grid_la_reg = expand.grid(alpha = 1,
			lambda = seq(0.001, 0.1, by = 0.0002))

# Training lasso regression model
lasso_model = train(x = deaths_train[,1:30],
					y = deaths_train$X31,
					method = "glmnet",
					trControl = control
					)
lasso_model

# mean validation score
lasso_rmse <- mean(lasso_model$resample$RMSE)

# Plot
plot(lasso_model, main = "Lasso Regression")
```

### Ridge

```{r}
library(lmridge)
```

```{r}
# build a ridge regression model and make prediction
ridge_model <- lmridge(X31 ~ ., deaths_train)
deaths_test$ri <- predict(ridge_model, deaths_test)
ri_rmse <- rmse(deaths_test$X31, deaths_test$ri)
```

### Elastic Net

```{r}
set.seed(42)
cv_5 = trainControl(method = "cv", number = 5)
```

```{r}
# build a elastic net
elnet = train(
  X31 ~ ., data = deaths_train,
  method = "glmnet",
  trControl = cv_5
)

elnet_int = train(
  X31 ~ . ^ 2, data = deaths_train,
  method = "glmnet",
  trControl = cv_5,
  tuneLength = 10
)
```

```{r}
# find out the best result
get_best_result = function(caret_fit) {
  best = which(rownames(caret_fit$results) == rownames(caret_fit$bestTune))
  best_result = caret_fit$results[best, ]
  rownames(best_result) = NULL
  best_result
}

result <- get_best_result(elnet_int)
en_rmse <- result$RMSE
```

### Result
```{r}
result <- data.frame(name = c("rf_rmse","gb_rmse","kn_rmse","lasso_rmse","ri_rmse","en_rmse"), 
                     rmse = c(rf_rmse,gb_rmse, kn_rmse,lasso_rmse,ri_rmse,en_rmse))
result[order(result[,1],decreasing=TRUE),]
```
Ridge regression model provides the smallest rmse 62.97624.

```{r}
# build a ridge regression model and make prediction for day one
ridge_model <- lmridge(X31 ~ ., deaths)
deaths$one <- predict(ridge_model, deaths)
deaths$one <- round(deaths$one)
for (i in 1:187){
  if (deaths$one[i] < deaths$X31[i]) {
    deaths$one[i] <- deaths$X31[i]
  }
}
```

```{r}
# day two
ridge_model <- lmridge(one ~ ., deaths)
deaths$two <- predict(ridge_model, deaths)
deaths$two <- round(deaths$two)
for (i in 1:187){
  if (deaths$two[i] < deaths$one[i]) {
    deaths$two[i] <- deaths$one[i]
  }
}
```

```{r}
# day three
ridge_model <- lmridge(two ~ ., deaths)
deaths$three <- predict(ridge_model, deaths)
deaths$three <- round(deaths$three)
for (i in 1:187){
  if (deaths$three[i] < deaths$two[i]) {
    deaths$three[i] <- deaths$two[i]
  }
}
```

```{r}
# day four
ridge_model <- lmridge(three ~ ., deaths)
deaths$four <- predict(ridge_model, deaths)
deaths$four <- round(deaths$four)
for (i in 1:187){
  if (deaths$four[i] < deaths$three[i]) {
    deaths$four[i] <- deaths$three[i]
  }
}
```

```{r}
# day five
ridge_model <- lmridge(four ~ ., deaths)
deaths$five <- predict(ridge_model, deaths)
deaths$five <- round(deaths$five)
for (i in 1:187){
  if (deaths$five[i] < deaths$four[i]) {
    deaths$five[i] <- deaths$four[i]
  }
}
```


```{r}
# Write data to txt file: tab separated values
# sep = "\t"
write.table(deaths, file = "deaths.txt", sep = "\t",
            row.names = TRUE, col.names = NA)
# Write data to csv files:  
# decimal point = "." and value separators = comma (",")
write.csv(deaths, file = "deaths.csv")
# Write data to csv files: 
# decimal point = comma (",") and value separators = semicolon (";")
write.csv2(deaths, file = "deaths.csv")
```

### Map 

```{r}
# load packages
library(janitor)
library(tidyr)
library(tidyverse)
```

```{r}
deaths[] <- lapply(deaths, function(x) type.convert(as.numeric(x)))
deaths$name <- rownames(deaths)
```

```{r}
joinData <- joinCountryData2Map( deaths,
                                 joinCode = "ISO3",
                                 nameJoinColumn = "name")
theMap <- mapCountryData( joinData, nameColumnToPlot="X31", addLegend=FALSE, mapTitle = "Current Deaths Cases")
do.call( addMapLegend, c(theMap, legendWidth=1, legendMar = 2))
```

```{r}
populationData <- deaths$one
joinData <- joinCountryData2Map( deaths,
                                 joinCode = "ISO3",
                                 nameJoinColumn = "name")
theMap <- mapCountryData( joinData, nameColumnToPlot="one", addLegend=FALSE, mapTitle = "Day 1 Forcast - Deaths Cases")
do.call( addMapLegend, c(theMap, legendWidth=1, legendMar = 2))
```

```{r}
populationData <- deaths$two
joinData <- joinCountryData2Map( deaths,
                                 joinCode = "ISO3",
                                 nameJoinColumn = "name")
theMap <- mapCountryData( joinData, nameColumnToPlot="two", addLegend=FALSE, mapTitle = "Day 2 Forcast - Deaths Cases")
do.call( addMapLegend, c(theMap, legendWidth=1, legendMar = 2))
```

```{r}
populationData <- deaths$three
joinData <- joinCountryData2Map( deaths,
                                 joinCode = "ISO3",
                                 nameJoinColumn = "name")
theMap <- mapCountryData( joinData, nameColumnToPlot="three", addLegend=FALSE, mapTitle = "Day 3 Forcast - Deaths Cases")
do.call( addMapLegend, c(theMap, legendWidth=1, legendMar = 2))
```

```{r}
populationData <- deaths$four
joinData <- joinCountryData2Map( deaths,
                                 joinCode = "ISO3",
                                 nameJoinColumn = "name")
theMap <- mapCountryData( joinData, nameColumnToPlot="four", addLegend=FALSE, mapTitle = "Day 4 Forcast - Deaths Cases")
do.call( addMapLegend, c(theMap, legendWidth=1, legendMar = 2))
```

```{r}
populationData <- deaths$five
joinData <- joinCountryData2Map( deaths,
                                 joinCode = "ISO3",
                                 nameJoinColumn = "name")
theMap <- mapCountryData( joinData, nameColumnToPlot="five", addLegend=FALSE, mapTitle = "Day 5 Forcast - Deaths Cases")
do.call( addMapLegend, c(theMap, legendWidth=1, legendMar = 2))
```

### end